---
title: "Final project"
author: "jk"
date: "2/22/2019"
output:
  pdf_document: default
  html_document: default
---
```{r}
airfoil<-read.table("/Users/joheunkang/Desktop/airfoil_self_noise.dat.txt")
library(car)
```

```{r}
freq <- airfoil$V1 #frequency
angle <- airfoil$V2 #frequency
chord <- airfoil$V3 #chord length 
vel <- airfoil$V4 #free stream velocity 
disp <- airfoil$V5 #displacement
sound <- airfoil$V6 #sound pressure level 
```

```{r}
#need to change vel and disp to categorical variables.
chord_lv<-as.factor(chord)
vel_lv<-as.factor(vel)
```

```{r}
mod.0<-lm(sound~1) #no model
mod.full<-lm(sound~ freq + angle + chord_lv + vel_lv + disp) #full model
pairs(sound~freq+angle+chord_lv+vel_lv +disp)
summary(mod.full) #summary of full model -> adj^R2 = 0.5214 , all predictors are important 
```
From pairs() angle and disp seems log relationship, freq and disp has inverse log relationship and angle and chorld_lv seems nagative relationship.\newline
From the Global F-test p-value < 2.2e-16 all five predictors are important.\newline
Adjusted R-squared: 0.5214
```{r}
library(leaps)
# finding a best model using forward method (AIC based).
forward_aic <-step(mod.0, scope = list(lower = mod.0,upper = mod.full), direction = "forward")
```
With AIC=4710.03, full model is the best model so far.

```{r}
# finding a best model using backward method (AIC based).
backward_aic <-step(mod.full, scope = list(lower = mod.0,upper = mod.full), direction = "backward")
```
full model is the best model so far.
```{r}
# finding a best model using stepwise method (AIC based)
stepwise_aic <- step(mod.0, scope = list(lower = mod.0,upper = mod.full))
```
full model is preferable.
```{r}
# finding a best model using forward method (BIC based).
# #observations = 1053
forward_bic<-step(mod.0, scope = list(lower = mod.0,upper = mod.full), direction = "forward", k = log(1053))
```
full model is preferabel.
```{r}
# finding a best model using backward method (BIC based)
backward_bic<-step(mod.0, scope = list(lower = mod.0,upper = mod.full), direction = "backward", k = log(1053))
backward_bic
```
????????????
```{r}
# finding a best model using stepwise method (BIC based)
stepwise_bic <- step(mod.0, scope = list(lower = mod.0,upper = mod.full),k = log(1053),trace = 0)
stepwise_bic
```
```{r}
#regsubet()
library(leaps)
mod.reg<-regsubsets(cbind(freq,angle,chord_lv,vel_lv,disp), sound, data = airfoil)
summary.reg<- summary(mod.reg)
names(summary.reg)
```

```{r}
# get all possible models from our full model.
summary.reg$which
```
we observed that the freq might be the most important predictor in our model.
```{r}
summary.reg$adjr2
```
Full model (model 5) has the biggest adjr2 0.5177846, one can consider our full model as the best model.

```{r}
summary.reg$cp
```
Since all other reduced model has Cp much bigger than q = (1+p), one consider out full model as the best model.

```{r}
summary.reg$bic
```
Since full model has significantly lower bic then all other models, one can consider full model as a best model. 
```{r}
summary.reg$rss
```
Full model always has lowest RSS so we normally choose the reduced model that has least rss. However, since there is significant difference between model 4 and full model (35654.58-34419.99 = 1234.59), one can choose full model as a best model. 
```{r}
# mod.full= lm(sound~freq+angle+chorld_lv+vel_lv +disp)
#compare mod.red1 and full model

mod.red1<-lm(sound~freq + angle)
anova(mod.red1,mod.full)
```
p-value:2.2e-16 < 0.05, at least one of chord_lv, vel_lv, disp is usful predictor for estimating sound.
F-value:103.8 >> 8
Conclusion: prefer to use Model 2 (full model).
```{r}
#compare mod.red2 and full model
mod.red2<-lm(sound~freq + angle+ chord_lv +vel_lv)
anova(mod.red2,mod.full)
```
p-value:2.375e-12 < 0.05, disp is useful important predictor for estimating sound.\newline
F-value:49.983 >> 8\newline
Conclusion: prefer to use Model 2 (full model).\newline\newline

We conclude that all five predictors are important for estimating sound.\newline

```{r}
#Residual-Fitted plot
plot(mod.full,which = 1)
```
Violations: Linearlity (not evenly scattered), Equal variances (pattern exists).\newline

```{r}
#Q-Q plot
plot(mod.full, which = 2)
```
Most of the points are on the Q-Q line, one can conlcuse that the residuals (?) are normally distributed.\newline
```{r}
# noise on angle
median(angle)
angle.mod <- angle +5.4
lm.angle.mod<-lm(sound~freq + angle+ chord_lv +vel_lv +disp)
summary(lm.angle.mod)
```
Since angle and many zeros, we add median (angle)  = 5.4 \newline
?? However, since there is no difference on summary table we keep original angle. 
```{r}
# trnasformation on predictors using powerTansform()
trans.mod <- powerTransform(cbind(freq,angle.mod,disp)~1, airfoil)
summary(trans.mod)
```
$\textbf{Rounded power}$:freq = 0.00, angle.mod = 0.13 and disp =  0.05. For sake of simplicity, we use $\lambda = 0$ for all of predictors.\newline
$\textbf{LR test, lambda = (0 0 0)}$: p-value: 0.00059349. Therefore, we do not take log transformation for all those three predictors. There exists at least one predictor that does not need log transformation.\newline
$\textbf{LR test, lambda = (1 1 1)}$: p-value:< 2.22e-16. Therefore, one can conclude that  exists at least one predictor that needs log transformation.\newline
```{r}
# trial 1
#log-transformation on disp
logtrans.disp.lm<-lm(sound ~ freq + angle + chord_lv + vel_lv + log(disp),data = airfoil)
plot(logtrans.disp.lm)
summary(logtrans.disp.lm)
```
```{r}
#transform on sound
boxCox(logtrans.disp.lm)
box.logtrans.disp.lm<-lm(sound**2~ freq + angle + chord_lv + vel_lv + log(disp),data = airfoil)
plot(box.logtrans.disp.lm)
summary(box.logtrans.disp.lm)
```
Adjusted R-squared:  0.5046\newline
After transform sound with $\lambda = 2$, we get $\textbf{lower Adjusted R-squared}$. So we decide not to transform on sound.

```{r}
# trial 2. 
# log-transformation on freq and angle+5.4
logtrans.freq.angle.lm<-lm(sound ~ log(freq) + log(angle.mod) + chord_lv + vel_lv + disp ,data = airfoil)
plot(logtrans.freq.angle.lm)
summary(logtrans.freq.angle.lm)
```
After log-transformaiton on bith freq and angle, we observed that linearlity and equal varieance have been worse. In this case, we do not need to try transformation on sound. 
```{r}
#trial 3
logtrans.angle.lm<-lm(sound ~ freq + log(angle.mod) + chord_lv + vel_lv + disp ,data = airfoil)
plot(logtrans.angle.lm)
summary(logtrans.angle.lm)
```

```{r}
#transform on sound
boxCox(logtrans.angle.lm)
```
Since $\lambda = 1$, we do not need to transform sound.\newline\newline
Without any interaction term, we consider our best model as \newline
sound ~ freq + angle + chord_lv + vel_lv + log(disp)\newline
$\textbf{Adding interaction term(s)}$
```{r}
#pairs() to check possibility of interaction terms
pairs(sound ~ freq + angle + chord_lv + vel_lv + log(disp))
```
From the scatter plot, one can observe that there might be a correlation between $\textbf{angle and log(disp)}$ and
chord_lv and log(disp).\newline\newline
$\textbf{Interaction Term}$
```{r}
#adding interaction term angle*log(disp)
int.1<-lm(sound ~ freq + angle + chord_lv + vel_lv + log(disp) + angle*log(disp))
plot(int.1)
summary(int.1)
```
By adding an interaction term angle * log(disp), Adjusted R-squared changed to 0.5158 which is 0.1% greater than the model without interaction term\newline
```{r}
#adding interaction term chord_lv*log(disp)
int.2<-lm(sound ~ freq + angle + chord_lv + vel_lv + log(disp) + chord_lv*log(disp))
plot(int.2)
summary(int.2)
```
By adding an interaction term chord_lv*log(disp), Adjusted R-squared changed to 0.5218  which is 1% greater than the model without interaction term\newline\newline
From adding each interaction term, one can observe some improvements on Adjusted R-squared and residual-fitted. However, evaluating the tradeoff between goodness of fit and model complexity one can say that the model without interaction term is more efficientway to estimate sound.\newline

$\textbf{Influential Points}$: 
```{r}
# full model Adjr2 = 0.5124 , log-tansfrom on disp Adjr2 =  0.5126 

#find influention points
logtrans.disp.lm<-lm(sound ~ freq + angle + chord_lv + vel_lv + log(disp),data = airfoil)
plot(logtrans.disp.lm)
```
From the Residual- Fitted plot, we observed thtat 1165,1045,1031 are the strong candidates for influential points. However, we need to check both cooks distance and high leverge test for the three points.
```{r}
# Test for outliers
outlierTest(logtrans.disp.lm)
```
Form the outlierTest, we observed that 1031 has studentized residual $|-3.822896|$ > 2. Therefore we conclude that 1031 is an outlier. 	
```{r}
# Look at leverage and Cook's distance
n = 1053
ft.cooks <- cooks.distance(logtrans.disp.lm)
which(ft.cooks > 4/(n-11-1))
```
```{r}
influenceIndexPlot(logtrans.disp.lm, id=TRUE)
```
Diagnostic Plots, we observed that 1030 has both hat-value and Cook's distance significantly big and 1217 has biggest cook's distane. This means that there will be relatively sigificant changes if we remove those two points from our data set. Therefore, we decide to remove those two points from our data set.
```{r}
# removing the influential points and define a new data set "airfoilclean"
airfoilclean <- airfoil[-c(1030,1217),] 

#resetting the variables 
n_freq <- airfoilclean$V1 #frequency
n_angle <- airfoilclean$V2 #frequency
n_chord <- airfoilclean$V3 #chord length 
n_vel <- airfoilclean$V4 #free stream velocity 
n_disp <- airfoilclean$V5 #displacement
n_sound <- airfoilclean$V6 #sound pressure level (RESPONSE)
n_angle.mod <- n_angle+5.4
```

```{r}
n_chord_lv <- as.factor(n_chord)
n_vel_lv <- as.factor(n_vel)
```

```{r}
n.mod.full <- lm(n_sound ~ n_freq + n_angle + n_chord_lv + n_vel_lv + log(n_disp))
plot(n.mod.full)
summary(n.mod.full)
ncvTest(n.mod.full)
```
Removing the observations 1030 and 1217, significant code for angle decreases from 0.01707 to 0.00656 and Adjusted R-squared went up 51.36% to 51.58%. From these changes, we discovered that these two points are indeed “influential” and, thus,it is crucial to exclude them from our dataset.\newline
However, p-value from non-constant variance test,1.1958e-07 is still less than 0.05.(non-equal variances)\newline\newline
$\textbf{(WLS) method}$
```{r}
# 1  Weight = 1/n.mod.full$fitted.values
n.mod.full.w1 <- lm(n_sound ~ n_freq + n_angle + n_chord_lv + n_vel_lv + log(n_disp),weights = 1/n.mod.full$fitted.values)
plot(n.mod.full.w1)
ncvTest(n.mod.full.w1)
```

```{r}
#2  Weight = 1/n_freq
n.mod.full.w2 <- lm(n_sound ~ n_freq + n_angle + n_chord_lv + n_vel_lv + log(n_disp),weights = 1/n_freq)
ncvTest(n.mod.full.w2)
```